---
title: "Basic Inferential Data Analysis"
author: "Miguel Angel Huerta"
date: "12 de marzo de 2018"
output:
  #pdf_document: default
  html_document: default
  word_document: default
---

##Part 2: Basic Inferential Data Analysis Instructions

Now in the second portion of the project, we're going to analyze the ToothGrowth data in the R datasets package.

  1.  Load the ToothGrowth data and perform some basic exploratory data analyses
  2.  Provide a basic summary of the data.
  3.  Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
  4.  State your conclusions and the assumptions needed for your conclusions.

1.- Exploratory Analysis

```{r Inf1}
library(ggplot2)
# Load ToothGrowth data
data("ToothGrowth")
head(ToothGrowth)
str(ToothGrowth)
summary(ToothGrowth)

ggplot(ToothGrowth, aes(x=factor(dose), y=len, fill=supp)) +
  geom_bar(stat="identity") +
  facet_grid(. ~ supp) +
  xlab("Dose (mg/day)") +
  ylab("Tooth length") +
  ggtitle("Tooth Length vs. Dose by Supplement")

ggplot(aes(x=dose, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=factor(dose)))+ facet_grid(~ supp)+ xlab("Supplement") + ylab("Tooth Length") + ggtitle("Tooth Length vs. Dose by Supplement")+   theme(plot.title = element_text(lineheight=.8, face="bold"))

ggplot(ToothGrowth, aes(x=factor(supp), y=len, fill=dose)) +
  geom_bar(stat="identity") +
  facet_grid(. ~ dose) +
  xlab("Dose (mg/day)") +
  ylab("Tooth length") +
  ggtitle("Tooth Length vs. Supplement  by Dose Amount")

ggplot(aes(x=supp, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=supp))+ facet_grid(~ dose)+ xlab("Supplement") + ylab("Tooth Length") + ggtitle("Tooth Length vs. Supplement by Dose Amount")+   theme(plot.title = element_text(lineheight=.8, face="bold"))


```

2.- Hyphotesis Testing

Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)

```{r Inf2}
# Supplement
h1 <- t.test(len~supp,data=ToothGrowth)
h1
# Supplements with Dose amounts 0.5
h2 <- t.test(len ~ supp, data = subset(ToothGrowth, dose == 0.5))
h2
# Supplements with Dose amounts 1
h3 <- t.test(len ~ supp, data = subset(ToothGrowth, dose == 1))
h3
# Supplements with Dose amounts 2
h4 <- t.test(len ~ supp, data = subset(ToothGrowth, dose == 2))
h4
```

3.- Conclusions
State your conclusions and the assumptions needed for your conclusions.

**Hypothesis1: One supplement delivers more tooth growth than the other**
```{r conc1}
h1$conf.int

paste0("The P value ",h1$p.value,"> 0.05, we cannot reject the null hypothesis since there is no statisical significance. ")
```

**REJECTED**

**Hypothesis2: One supplement delivers more tooth growth than the other at dose 0.5**
```{r conc2}
h2$conf.int
h2$p.value
paste0("The P value ",h2$p.value,"< 0.05, the difference is statistically significant so we can reject the null hypothesis.The alternative hypothesis that one supplement delivers more tooth growth than the other at  0.5 mg/day  is accepted. ")
```
**ACCEPTED**

**Hypothesis3: One supplement delivers more tooth growth than the other at dose 1**
```{r conc3}
h3$conf.int
h3$p.value
paste0("The P value ",h3$p.value,"< 0.05, the difference is statistically significant so we can reject the null hypothesis.The alternative hypothesis that one supplement delivers more tooth growth than the other at  1 mg/day  is accepted. ")
```
**ACCEPTED**

**Hypothesis4: One supplement delivers more tooth growth  at dose 2**
```{r conc4}
h4$conf.int
h4$p.value
paste0("The P value ",h1$p.value,"> 0.05, we cannot reject the null hypothesis since there is no statisical significance. ")
```
**REJECTED**

**FINAL CONCLUSION**

We can ensure with less than a %5 of coincidence that at doses of 0.5 mg/day and 1mg/day the tooth length is greater with OJ supplement- than with VC.


