---
title: "Simultation Example"
author: "Miguel Angel Huerta"
date: "6 de marzo de 2018"
#output:
  #word_document: default
  #pdf_document: default
  #html_document: default
output: 
       rmdformats::readthedown
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```


##  Simulation Exercise Instructions 

In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

1.- Show the sample mean and compare it to the theoretical mean of the distribution.

```{r simmulation}
library(ggplot2)
n=40
lambda =.2
mu <- 1/lambda
s <- 1/lambda
sim <- 1000 

obs <- rexp(n,lambda)

plot(obs, pch=20, cex=0.6, main=paste0("The exponential distribution with lambda ",lambda," and ",n," observations"))
abline(h=mean(obs), col="red", lwd=3, lty=2)
text(100, 30, paste("Actual mean = ", round(mean(obs),2), "\n Theoretical mean = ",mu ), col="red")
# theoretical mean  vs sample mean
print (paste("Population mean: ", mu, ", Sample mean: ", round(mean(obs) ,4) ) )

hist(rexp(n,lambda), main=paste0("Histogram of ",n," exponential observations"))
abline(v=mean(obs), col="red", lwd=3, lty=2)
abline(v=mu, col="green", lwd=3, lty=2)
text(10,10, paste("Actual mean = ", round(mean(obs),2)), col="red")
text(10,9, paste(" ", "\n Theoretical mean = ",mu ), col="green")

```
```{r simmulation2}
library(ggplot2)
sample_size=40
lambda =.2
mu <- 1/lambda
s <- 1/lambda
num_sim <- 1000 

observations <- matrix(rexp(num_sim * sample_size, rate = lambda), num_sim, sample_size)
means <- rowMeans(observations)

plot(observations, pch=20, cex=0.6, main=paste0("Exponential distribution with lambda ",lambda,",sample size of ",sample_size," and ",num_sim," simulations"))
abline(h=mean(observations), col="red", lwd=3, lty=1)
abline(h=mu, col="green", lwd=3, lty=2)
text(25, 30, paste("Actual mean = ", round(mean(observations),2)), col="red")
text(25, 28, paste("Theoretical mean = ",mu ), col="green")
# theoretical mean  vs sample mean
print (paste("Population mean: ", mu, ", Sample mean: ", round(mean(observations) ,4) ) )

hist(observations, main=paste0("Histogram of ",n," exponential observations and ",sim," simulations"))
abline(v=mean(observations), col="red", lwd=3, lty=1)
abline(v=mu, col="green", lwd=3, lty=2)
text(20,10000, paste("Actual mean = ", round(mean(observations),2)), col="red")
text(20,9000, paste("Theoretical mean = ",mu ), col="green")


```
2.- Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

```{r simm3}
# theoretical standard deviation vs practical standard deviation
print (paste("Theoretical standard deviation: ", round( (1/lambda)/sqrt(n) ,4), ", Practical standard deviation", round(sd(means) ,4) ) )


# the variance:
print (paste("Theoretical variance: ", (1/lambda)^2/n, ", Practical variance", round(var(means) ,4) ) )
```


3.- Show 1000 simulaton histogram

```{r simm2}
library(ggplot2)
means<-matrix(ncol=1,nrow = 1000)
for (i in 1:sim){
 means[i,]<- matrix(mean(rexp(n,lambda)))
}

hist(means)
abline(v=mean(means), col="red", lwd=3, lty=2)
abline(v=mu, col="green", lwd=3, lty=2)
text(3.6,90, paste("Actual mean = ", round(mean(means),2), "\n Theoretical mean = ",mu ), col="red")


```

4.- The sample means distribution follows a teorical normal distribution

```{r simm4}
#df <- data.frame(x = rnorm(1000, mu))
df <- data.frame(x = means)

ggplot(df, aes(x)) +
  geom_histogram(aes(y = ..density..)) +
  stat_function(fun = dnorm, args = list(mean = mu,sd = s^2/n),lwd = 2, col = 'green')
```

5.- Final plot with all the information

```{r final}

library(ggplot2)
sample_size=40
lambda =.2
mu <- 1/lambda
s <- 1/lambda
num_sim <- 1000 

observations <- matrix(rexp(num_sim * sample_size, rate = lambda), num_sim, sample_size)
means <- rowMeans(observations)

hist(means, prob = TRUE, main = "Mean Distribution for 40 Exponential observations with 1000 simulations", xlab = "Simulation Mean")
lines(density(means),col = "red")
abline(v=mean(means), col="red", lwd=3, lty=1)
abline(v=mu, col="green", lwd=3, lty=2)
x_sim <- seq(min(means), max(means), length = 100)
dist_sim <- dnorm(x_sim, mean = 1/lambda, sd = (1/lambda/sqrt(sample_size)))
lines(x_sim, dist_sim, pch = 50, col = "green", lty = 5)
legend('topright', c(paste0("Simulation(",round(mean(means),2),")"), paste0("Theoretical Normal(",mu,")")), 
       lty = c(1,5), col = c("red", "green"))
```
The comparison of the sample distribution with the theoretical normal distribution will be done in the following three ways.

1) we visually compare the graph of the distributions. We can see that the distribution of sample means (the histogram) closely matches a theoretical normal distribution.

2) we need to compare the mean, variance, and confidence intervals between two distributions. In the above sections titled Distribution - Sample Mean vs. Theoretical Mean and Distribution - Sample Variance vs. Theoretical Variance, we compared the mean and variance and showed that they closely match. Now, the confidence intervals:

```{r IC}
simul_interval <- round(mean(means) + c(-1,1)*1.96*sd(means)/sqrt(sample_size),3)
theo_interval <- mu + c(-1,1)*1.96*sqrt(s)/sqrt(sample_size)
```

The 95% confidence interval for the simulation is (4.763, 5.26) while the theoretical 95% confidence interval is (4.755, 5.245). We can conclude that the confidence intervals also closely match each other.

3) we look at the QQ plot, wich also show very similar curve for sample quantiles and theorical quantiles.

```{r QQPLOT}
qqnorm(means)
qqline(means)
```

