---
title: "Storm"
author: "Miguel Angel Huerta"
date: "15 April  2018"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(RCurl)
library(ggplot2)
library(data.table)
library(Hmisc)
library(dplyr)

if (!file.exists("data")){
  dir.create("data")
  fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
  con <- download.file(fileUrl,destfile = "./data/activity.bz2",method = "libcurl")
}

storm <- read.csv(file = "./data/activity.bz2", sep=",", header=T)
event <- c("EVTYPE", "FATALITIES", "INJURIES", "PROPDMG", "PROPDMGEXP", "CROPDMG", 
           "CROPDMGEXP")
data <- storm[event]
```

#Synopsis

In the present study i am going to make use of public NOAA Storm Database to show the events that are  most harmful with respect to population health and which types of events have the greatest economic consequences.

#Data Analysis

```{r dataAnalysis, echo=TRUE}


unique(data$PROPDMGEXP)
# Assigning values for the property exponent data 
data$PROPEXP[data$PROPDMGEXP == "K"] <- 1000
data$PROPEXP[data$PROPDMGEXP == "M"] <- 1e+06
data$PROPEXP[data$PROPDMGEXP == ""] <- 1
data$PROPEXP[data$PROPDMGEXP == "B"] <- 1e+09
data$PROPEXP[data$PROPDMGEXP == "m"] <- 1e+06
data$PROPEXP[data$PROPDMGEXP == "0"] <- 1
data$PROPEXP[data$PROPDMGEXP == "5"] <- 1e+05
data$PROPEXP[data$PROPDMGEXP == "6"] <- 1e+06
data$PROPEXP[data$PROPDMGEXP == "4"] <- 10000
data$PROPEXP[data$PROPDMGEXP == "2"] <- 100
data$PROPEXP[data$PROPDMGEXP == "3"] <- 1000
data$PROPEXP[data$PROPDMGEXP == "h"] <- 100
data$PROPEXP[data$PROPDMGEXP == "7"] <- 1e+07
data$PROPEXP[data$PROPDMGEXP == "H"] <- 100
data$PROPEXP[data$PROPDMGEXP == "1"] <- 10
data$PROPEXP[data$PROPDMGEXP == "8"] <- 1e+08
# Assigning '0' to invalid exponent data
data$PROPEXP[data$PROPDMGEXP == "+"] <- 0
data$PROPEXP[data$PROPDMGEXP == "-"] <- 0
data$PROPEXP[data$PROPDMGEXP == "?"] <- 0
# Calculating the property damage value
data$PROPDMGTOT <- data$PROPDMG * data$PROPEXP

unique(data$CROPDMGEXP)

# Assigning values for the crop exponent data 
data$CROPEXP[data$CROPDMGEXP == "M"] <- 1e+06
data$CROPEXP[data$CROPDMGEXP == "K"] <- 1000
data$CROPEXP[data$CROPDMGEXP == "m"] <- 1e+06
data$CROPEXP[data$CROPDMGEXP == "B"] <- 1e+09
data$CROPEXP[data$CROPDMGEXP == "0"] <- 1
data$CROPEXP[data$CROPDMGEXP == "k"] <- 1000
data$CROPEXP[data$CROPDMGEXP == "2"] <- 100
data$CROPEXP[data$CROPDMGEXP == ""] <- 1
# Assigning '0' to invalid exponent data
data$CROPEXP[data$CROPDMGEXP == "?"] <- 0
# calculating the crop damage value
data$CROPDMGTOT <- data$CROPDMG * data$CROPEXP


DT <- data.table(data)

PROP <- DT[,sum(PROPDMGTOT),by=EVTYPE]
PROP <- PROP[order(-PROP$V1),]
PROP <- PROP[1:5,]
PROP$V1 <- PROP$V1/1000000

CROP <- DT[,sum(CROPDMGTOT),by=EVTYPE]
CROP <- CROP[order(-CROP$V1),]
CROP <- CROP[1:5,]
CROP$V1 <- CROP$V1/1000000

```

#Results

Here we can see Drought is the  most harmful with respect to population health. Here are the top 5.
```{r result, echo=TRUE}
    ggplot(data=CROP, aes(x=CROP$EVTYPE,y=CROP$V1))+
      geom_bar(stat="identity") +
      xlab("Event types") +
      ylab("Millions fatalities") +
      theme_minimal()
```

And TSTM WIND have the greatest economic consequences. Here are the top 5.
```{r resultPROP, echo=TRUE}
    ggplot(data=PROP, aes(x=PROP$EVTYPE,y=PROP$V1) )+
      geom_bar(stat="identity") +
      xlab("Event types") +
      ylab("Millions $") +
      theme_minimal()

```